---
title: "Price Tracker - Shiny App"
author: "28Melvin"
date: "2021/02/01"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Melvin There!

Github: Yet to come\
Shinyapps.io: [Price Tracker ShinyApp](https://melvin28.shinyapps.io/pricetracker) 

This is/was basically my first RShiny project.So what’s supposed to happen?

The Shiny App offers the user the possibilty to extract the price of either the “IPhone 12 Pro - 256 GB” or “Air Pods mit Ladecase” from the Apple Website (German Site - € Prices) The user can then choose to plot the price development of either product.

General Structure: 

2 .R files
   - app.R
   - global.R

1 .Renviron file
  - .Renviron
  
All three files are uploaded to shinyapps.io where the app is pusblished. Data is stored in a remote MySQL database. 

It’s my first app and thus lot’s of issues need to be improved. - prices on the apple website hardly vary over time, thus getting prices from different site might be more interesting - data should only be written to the database once a day, thus not always when a user decides to extract the new price (otherwise there would be quite a lot entries/duplicates in the database if the app WOULD be frequently used)

# The global.R file

As far as I'm concerned when deploying the Shiny App on Shinyapps.io, the global.R file is automatically executed when the app is launched. However only if the "user interface" and the server function are in two separated files. When deploying in a single app.R file, there needs to be the source-Function included. 


```{r, eval=FALSE}

# load libraries
library(xml2)  
library(rvest)
library(stringr)
library(lubridate)

library(ggplot2)

library(RMySQL)


# get the prices 
# two functions - 1. IPhone, 2.AirPods 


getIPhone <- function(){
  
  getSite_iphone <- read_html("https://www.apple.com/de/shop/buy-iphone/iphone-12-pro/6,1%22-display-256gb-graphit")
  
  getPrice <- getSite_iphone %>%
    html_nodes("span.as-price-currentprice") %>%  #span = tag name, as-price-currentprice = class name
    html_text()
  getPrice <- str_remove_all(getPrice, "[\n]") 
  getPrice <- gsub(" ","", getPrice)
  getPrice <- gsub(".","",getPrice, fixed=TRUE) #fixed=TRUE so that only the "." is replaced
  getPrice <- str_replace_all(getPrice, ",\\w+", "") #remove all character after "," included, so that also the "€" sign is removed
  getPrice <- substr(getPrice, 1,nchar(getPrice)-1)
  getPrice_iphone <- as.integer(getPrice)
  
  
  
  return(getPrice_iphone)

}



getAirPod <- function(){
  
  getSite_airpod <- read_html("https://www.apple.com/de/shop/product/MV7N2ZM/A/airpods-mit-ladecase")
  
  getPrice <- getSite_airpod %>%
    html_nodes("span.current_price") %>%
    html_text()
  
  
  getPrice <- str_remove_all(getPrice, "[\n]") 
  getPrice <- gsub(" ","", getPrice)
  getPrice <- str_replace_all(getPrice, ",\\w+", "") #remove all character after "," included
  getPrice <- substr(getPrice, 1,nchar(getPrice)-1)
  getPrice_airpod <- as.integer(getPrice)
  
  return(getPrice_airpod)
  
}

# the function only creates a data.frame that stores the current date, as well as the year
# later these two values are combined with product name and the price of the product into a data.frame that is finally store in the MySQL database
getTime <- function(){
  
  
  YearMonthDate <- Sys.Date()
  Year <- year(YearMonthDate)
  inputtable <- as.data.frame(cbind(Year,YearMonthDate))
  inputtable$YearMonthDate <-  as.numeric(inputtable$YearMonthDate)
  inputtable$YearMonthDate <- format(as.Date(inputtable$YearMonthDate, origin="1970-01-01"), "%Y:%m:%d")
  
  return(inputtable)
  
  
}


# Database access (save to database and load from database)

table <- "PriceTracker" #name of the table in the database 



saveData <- function(data){
  
  #connect to database
  
  db <- dbConnect(MySQL(), dbname = Sys.getenv("databaseName"),  
                  host = Sys.getenv("host"),  
                  port = as.integer(Sys.getenv("port")), 
                  user = Sys.getenv("user"),  
                  password = Sys.getenv("password"))
  
  #######################################################################################
  # The database credentials are stored in a .Renviron file that was also deployed to 
  # the shinyapps.io server. 
  # When testing locally the .Renviron file was project specific -> usethis::edit_r_environ("project") 
  ##################################
  
  
  # construct update query by looping over the data fields
  
  query <- sprintf(
    "INSERT INTO %s (%s) VALUES ('%s')", 
    
    table, # 
    paste(names(data), collapse = ", "),
    paste(data, collapse = "', '")    
    
  )
  
  # Submit the update query and disconnect
  
  dbGetQuery(db, query)
  
  dbDisconnect(db)
  
}


loadData <- function(){
  
  # connect to the database
  db <- dbConnect(MySQL(), dbname = Sys.getenv("databaseName"), 
                  host = Sys.getenv("host"),  
                  port = as.integer(Sys.getenv("port")), 
                  user =  Sys.getenv("user"), 
                  password = Sys.getenv("password")) 
  
  # construct the fetching query 
  
  query <- sprintf("SELECT * FROM %s", table)
  
  # Submit the fetch query and disconnect
  
  data <- dbGetQuery(db, query)
  
  dbDisconnect(db)
  
  
  return(data)
  
}



# the plot functions - these are later executed once the user in the app hits the "plot button"
# the data input for the plot functions is the output of the loadData function / the data from the MySQL database 

plotAirPods <- function(input){
  
  
  input1 <- input[which(input[,"ProductName"] == "AirPods mit Ladecase"),]
  #prepare data
  ggplot(data = input1, aes(x = YearMonthDate, y = Price, group = 1)) + geom_line()
  
  
}

plotIPhone <- function(input){
  
  input1 <- input[which(input[,"ProductName"] == "Apple IPhone 12 Pro 256 GB"),]
  
  ggplot(data = input1, aes(x = YearMonthDate, y = Price, group = 1)) + geom_line()
  
}


```

# The app.R File


```{r, eval=FALSE}
# r-shiny project
library(shiny) 

source("./global.R") #relative path as otherwise the file cannot be found when deploying to shinyapps.io


# Define UI / Layout 
# fluidpage -> function
ui <- fluidPage(
  
  titlePanel(""),
  

  
  sidebarLayout(
    
    sidebarPanel("Choose the product you wish to get the price for: ", 
                 
                 br(),
                 br(),
                 selectInput("productID","Choose product: ", c("Apple IPhone 12 Pro 256 GB", "AirPods mit Ladecase")),
                 actionButton("getNewPrice", "Load current Price"),
                 br(),
                 br(),
                 actionButton("plotData", "Plot Data"),
                 br(), 
                 h1(textOutput("currentPrice"), align = "center")), 
                 
               
    
    mainPanel(p("Prices (€) over Time", align = "center"), tabsetPanel(type = "tabs", 
                          tabPanel("IPhone 12 Pro - 256 GB", plotOutput("plotIPhone")),
                          tabPanel("AirPods mit Ladecase", plotOutput("plotAirPods"))),
              
              position = "left" #sidebar position
              
         
        
  ))
  
)


# Define server logic required 
server <- function(input, output) {
  
  tableInput <- loadData() #as to my knowledge loading the data at this point, loads the data everytime a user joins the app
                          #I got to be honest, I don't know if it is better to place it in the global.R app, that is supposed to only run once the app launches (as I have no seperate ui.R and server.R files)
  
  # if the user pushes the "Get newest Price" Button - newest price is extracted 
  # and written to remote database + current price is displayed (however without € sing. 
  observeEvent(input$plotData, {
    
    
    
    
    if(input$productID == "Apple IPhone 12 Pro 256 GB"){
      
      output$plotIPhone <- renderPlot(plotIPhone(tableInput))
      
    }
    else
    {
    
     output$plotAirPods <- renderPlot(plotAirPods(tableInput))
    }
    
   # output$datas <- renderTable(tableInput)
    
    
  })
  
  
  
  observeEvent(input$getNewPrice, {
    
    
    # first we scrape the price from the homepage
    # thereafter we get the current date and combine these two inputs
    # the data frame that is then written to the remote database
    
    if(input$productID == "Apple IPhone 12 Pro 256 GB"){
      
      Price <- getIPhone() # scrape IPhone Price
      
    }else{
      
      Price <- getAirPod()
    }
    
    times <- getTime() 
 
    ProductName <- input$productID
    all_frame <- as.data.frame(cbind(ProductName, Price, times))
    
    output$currentPrice <- renderText({
      
      print(Price)
      
    })
    
    saveData(all_frame)
    
    
  })
  
  
  
 
  
}


# run the app 
shinyApp(ui = ui, server = server)

```

# The .Renviron File

The following should be included in the .Renviron file. The .Renviron file should also be uploaded to the shinyapps.io server. 
When testing the app locally I used the: usethis::edit_r_environ("projects") command to create the .Renviron file. 
```{r, eval= FALSE}


host = put_the_host_here
port = put_port_here
user = put_user_here
password = put_password_here
databaseName = put_databaseName_here
table = put_tableName_here

```
